Canonical rings of $\bq$-divisors on Minimal Rational Surfaces

### 2015-07-27 surfaces-write-up/surface-write-up.tex
some edits to relations lemma for hirz

\begin{defn}\label{def:ell_ij}
Let $D = \sum_{i=1}^n \alpha_i D_i$ where $\alpha_i = \frac{c_i}{k_i}$.
\todo{Peter: better notation for fraction? Aaron: I like this notation}  Then notate
\[\ell_{i,j} = \lcm_{h\ne i,j}(k_h).\]
\end{defn}

@@

Let $\omega\in \Omega'$ correspond to the monomial $y_i \prod_{j=1}^n
{y_j}^{c_j}$.  Then we can write $\omega$ as a sum of $r_j e_j$'s for $j\in
\Te$ and $r_{j,k}e_{j,k}$ for $j\in \Tp, k\in \Tm$.  
For convenience, let $e_j$ (when it exists) be in degree $d_j$ with $j^{th}$
component $-\alpha_j d_j + \kappa_j$ for some $\kappa_j \in \bq$ and let $e_{j,k}$ (when it exists) lie in
degrees $d_{j,k}$ with $j^{th}$ component $-\alpha_j d_{j,k} + \kappa_{j,k}'$ and $k^{th}$
component $-\alpha_j d_{j,k} + \kappa_{j,k}''$ for $\kappa_{j,k}', \kappa_{j,k}'' \in \bq$
Then a careful computation deduces that

\todo{Aaron: What is $\sigma$ here? Is it $\rho$?}
%if $i\in \Te$, $i\in \Tp$, and $i\in \Tm$, then
\begin{equation}\label{eqn:sigma-in-terms-of-lambda/E}
	\sigma = \lambda + E
\end{equation}
where $\lambda \in \Sigma$ and
\begin{equation}\label{eqn:hirz-E-translation}
	E = \begin{cases}
	\frac{1}{\kappa_i} e_i  	&  \mbox{ if }i\in \Te\\
	\sum_{j\in \Tm} r_{i,j} e_{i,j}: r_{i,j}\in \bq, \sum_{j\in \Tm} r_{i,j}\kappa_{i,j}' = 1 & \mbox{ if } i\in \Tp\\
	\sum_{j\in \Tp} r_{j,i} e_{j,i}: r_{j,i} \in \bq, \sum_{j\in \Tm} r_{j,i} \kappa_{i,j}'' = 1 	& \mbox{ if } i\in \Tm\\
	\end{cases}.
\end{equation}

\noindent
From this, we can deduce that
\[
	\deg(E) \le \begin{cases}
	\ell_i \gcd(\bida_i, \bidb_i)	&\mbox{ if } i \in \Te \\
	\max_{j \in \Tm} \bigl(\ell_{i, j} (\bida_i \bidb_j - \bida_j \bidb_i)\bigr)
	&\mbox{ if } i \in \Tp \\
	\max_{j \in \Tp} \bigl(\ell_{j, i} (\bida_j \bidb_ i - \bida_i \bidb_j) \bigr)
	&\mbox{ if } i \in \Tm \end{cases}.
\]
Let $M$ be the set of monomials terms of $\beta_i$.  Then for $\mu\in M$, we
\todo{Aaron: What is $M$?}
\todo{Aaron: Again, why do we keep defining lattce points in
	terms of elements in the canonical ring? Why not just
write out what the lattice points are in coordinates?}
can further compute that the lattice point, $\sigma_\mu$, corresponding to $\mu :=
\prod_{j=1}^n
{f_i}^{c_i}$ is of the form
\[
	\sigma_\mu = \lambda + E_\mu
\]
where $\lambda \in \Sigma$ is the same as in Equation \ref{eqn:sigma-in-terms-of-lambda/E} and $E_\mu$ is defined in a similar fashion as $E$ in Equation \ref{eqn:hirz-E-translation} (in particular, they lie in the same degree).
\todo{Aaron: if $\sigma_\mu$ is an element, how can it be written as
	the sum of an element and a set? Same question for $\sigma$ above.
	Also, I do not know what $E_\mu$ is. You really have to define it,
since there's no way to check the validity of this proof without it.}
Let $F\in R_D$ corresponding to the lattice point $E$ and let $F_\mu$ correspond to the lattice point $E_\mu$.  
Then $F - \sum_{\mu \in M} F_\mu = 0$ induces a
relation in degree at most
\[
	\rho + \deg(F)
\]

>>>>

\begin{defn}\label{def:ell_ij}
Let $D = \sum_{i=1}^n \alpha_i D_i$ where $\alpha_i = \frac{c_i}{k_i}$.
Then notate
\[\ell_{i,j} = \lcm_{h\ne i,j}(k_h).\]
\end{defn}

@@

Let $\omega\in \Omega'$ correspond to the monomial $y_i \prod_{j=1}^n
{y_j}^{c_j}$.  Then we can write $\omega$ as a sum of $r_j e_j$'s for $j\in
\Te$ and $r_{j,k}e_{j,k}$ for $j\in \Tp, k\in \Tm$.  
For convenience, let $e_j$ (when it exists) be in degree $d_j$ with $j^{th}$
component $-\alpha_j d_j + \kappa_j$ for some $\kappa_j \in \bq$ and let $e_{j,k}$ (when it exists) lie in
degrees $d_{j,k}$ with $j^{th}$ component $-\alpha_j d_{j,k} + \kappa_{j,k}'$ and $k^{th}$
component $-\alpha_j d_{j,k} + \kappa_{j,k}''$ for $\kappa_{j,k}', \kappa_{j,k}'' \in \bq$
Then a careful computation deduces that we can write

\todo{Aaron: What is $\sigma$ here? Is it $\rho$? \\
Peter: Should have been $\omega$...}
%if $i\in \Te$, $i\in \Tp$, and $i\in \Tm$, then
\begin{equation}\label{eqn:sigma-in-terms-of-lambda/E}
	\omega = \lambda + E
\end{equation}
for some $\lambda \in \Sigma$ and $E$ of the form
\begin{equation}\label{eqn:hirz-E-translation}
	E = \begin{cases}
	\frac{1}{\kappa_i} e_i  	&  \mbox{ if }i\in \Te\\
	\sum_{j\in \Tm} r_{i,j} e_{i,j}: r_{i,j}\in \bq, \sum_{j\in \Tm} r_{i,j}\kappa_{i,j}' = 1 & \mbox{ if } i\in \Tp\\
	\sum_{j\in \Tp} r_{j,i} e_{j,i}: r_{j,i} \in \bq, \sum_{j\in \Tm} r_{j,i} \kappa_{i,j}'' = 1 	& \mbox{ if } i\in \Tm\\
	\end{cases}.
\end{equation}

\noindent
From this, we can further deduce that for such an $E$,
\[
	\deg(E) \le \begin{cases}
	\ell_i \gcd(\bida_i, \bidb_i)	&\mbox{ if } i \in \Te \\
	\max_{j \in \Tm} \bigl(\ell_{i, j} (\bida_i \bidb_j - \bida_j \bidb_i)\bigr)
	&\mbox{ if } i \in \Tp \\
	\max_{j \in \Tp} \bigl(\ell_{j, i} (\bida_j \bidb_ i - \bida_i \bidb_j) \bigr)
	&\mbox{ if } i \in \Tm \end{cases}.
\]
Let $M$ be the set of monomials terms of $\beta_i$.  Then for $\mu = \prod_{j=1}^n
{f_i}^{c_i}\in M$, we
\todo{Aaron: What is $M$? \\
Peter: I don't quite follow this question... Is your objection that ``the set of monomial terms of $\beta_i$'' is too imprecise?}
\todo{Aaron: Again, why do we keep defining lattce points in
	terms of elements in the canonical ring? Why not just
write out what the lattice points are in coordinates? \\ 
Peter: Fixed?}
can further compute that the lattice point, $\omega_\mu = (c_1 ,\ldots c_n)$, corresponding to $\mu
$ is of the form
\[
	\omega_\mu = \lambda + E_\mu
\]
where $\lambda \in \Sigma$ is the same as in Equation \ref{eqn:sigma-in-terms-of-lambda/E} and $E_\mu$ is defined in a similar fashion as $E$ in Equation \ref{eqn:hirz-E-translation}:
\begin{equation}\label{eqn:hirz-E_mu-translation}
E_\mu = \left(\sum_{\substack{i \in \{1, 2, 3, 4\} \cap \Te}}q_i e_i\right) + \left(\sum_{\substack{i \in \{1, 2, 3, 4\} \cap \Tp \\ j\in \Tm}} q_{i,j} e_{i,j}\right) + \left(\sum_{\substack{i \in \{1, 2, 3, 4\} \cap \Tm \\ j\in \Tp}} q_{j,i} e_{i,j}\right)
\end{equation}
for some $q_i$'s, $q_{i,j}$'s and $q_{j,i}$'s (for appropriate $i$'s and $j$'s) in $\bq_{\ge 0}$.  
\todo{Aaron: if $\sigma_\mu$ is an element, how can it be written as
	the sum of an element and a set? Same question for $\sigma$ above.
	Also, I do not know what $E_\mu$ is. You really have to define it,
since there's no way to check the validity of this proof without it.  \\
Peter: $E$ and $E_\mu$ are elements, not sets...}
Note that $E_\mu$ lies in the same degree as $E$ for all $\mu$.

Let $F\in R_D$ corresponding to the lattice point $E$ and let $F_\mu$ correspond to the lattice point $E_\mu$.  
Then $F - \sum_{\mu \in M} F_\mu = 0$ induces a
relation in degree at most
\[
	\rho + \deg(F)
\]
